---
output: html_document
---

# Introduction

## Scripts

R Code is executed line-by-line with `Ctrl + Enter`

```{r}
print("hello world")
```

Full scripts can be run with `Ctrl + Shift + Enter`

```{r}
stopifnot(TRUE)
```

## Vectors

R is built around vectors. (The most useful statistical object). Use the `c` (combine/concatenate) command to create a vector in R.

```{r}
v <- c(3,1,4,1,5,9)
```

Numbers are also (atomic) vectors. (Tip: `Alt + "-"` will generate the assignment notation.)

```{r}
n <- 110
```

There are many built-in functions, e.g., like `sum()` that adds all of the values in a vector. (Also: `max()`, `min()`, and `length()`.)

```{r}
sum(v)
```

A shortcut for creating a vector $1$ to $n$ is

```{r}
w <- 1:n
```

We can also define

```{r}
m <- 10; n <- 20
```

so that

```{r}
m:n
```

We can access the `i`th entry in a vector with `v[i]`, 
where `v[1]` is the first entry

```{r}
v[c(1,3,5)]
```

We can also create a sub-vector by deleting entries. 
Note, the brackets are needed because -2:4 generates -2, -1, 0,...

```{r}
v[-(2:4)]
```

Operations are interpreted component-wise. 
For example, the following command generates ($1/1, 1/2^2, \ldots$)

```{r}
head(1/(1:100)^2)
```

In mathematics, 
adding vectors `v + w` of different lengths is undefined. 
In R, recycling is used.

```{r}
head(v + w)
```
