---
output: html_document
---

# Events

## Sampling

The `sample` command can generate pseudo random numbers _without_ replacement

```{r}
n <- 10; k <- 5
sample(n, k)
```

Setting replace to "TRUE" for sampling _with_ replacement

```{r}
sample(n, k, replace = TRUE)
```

To generate a random permutation, use `sample(n, n)` or

```{r}
sample(n)
```

R can also draw from non-numeric vectors. 
For example

```{r}
sample(letters, 7)
```

## Naive Probability

The `sample` command, unless specified, 
will generate *equally likely* outcomes. 
We can check this by computing the long-run frequency of two outcomes, 
"heads" and "tails"

```{r}
set.seed(42)
sample_size <- 1e5
frequencies <- table(sample(c("heads", "tails"), sample_size, TRUE))
(probabilities <- frequencies / sample_size)
```

The probability of the event "heads", $P(A_H)$, is approximated by the long-run frequency

$$
\frac{n_H}{n} = 0.503 \approx 1/2 = \frac{|A_H|}{|S|} = P(A_H)
$$

The events "tails" and "heads" are approximately equal, confirming that each outcome is equally likely.
