\documentclass[11pt,a4paper]{article}

\usepackage{../notes}

\title{Probability Theory \& Statistics}

\begin{document}

\maketitle
\newpage

\section{Introduction}

\subsection{Motivation}

Data and statistics are the language of modern business and science. 
It is rare in the modern work place to embark on a task or project that doesn't require 
some form of collection of data and some form of description of that data. 
Probability and statistics are \emph{descriptions of data}. 
Moreover, 
the language they provide for describing data is rich and nuanced, 
but requires practice (as with any language be it German or Python!)

\subsection{Aim}

The sessions' aim to develop fluency in \emph{statistical language} and \emph{probabilistic thinking}.

\subsection{Delivery}

The sessions should occur every two weeks,
over the period of a year to two years depending on the depth of discussions during sessions.

\subsection{Fundamental Objects}

There are 4 fundamental objects%
\footnote{Thinking in terms of objects helps avoid \emph{category errors}.}
in probability theory:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item \emph{Events}
\item \emph{Random Variables}
\item \emph{Distributions}
\item \emph{Numbers}
\end{enumerate}

We will work through and discuss each of these objects 
and most importantly \emph{practice using them in problems}. 

\subsection{Structure}

\subsubsection{Prerequisites}

Each session tries to be self-contained, 
although this is sometimes difficult or impossible for certain topics, 
in which case a \emph{prerequisite} list of sessions or reading is suggested at the start of the session document. 
Further reading is also provided at the end of each session for a more in-depth discussion 
of the session topic.

\subsubsection{Notation}

New terms and concepts are \emph{emphasized}. 

\subsubsection{Interpretation}

We will try to avoid too much mathematical jargon and unnecessary proofs, 
keeping formal notation to the bare minimum and where necessary 
explained in plain language.

\subsubsection{Discussions}

Frequently asked questions from each session 
are added to the end of the session notes for future reference.

\subsubsection{Practice}

Practice questions and their solutions are provided, 
the expectation is that they are worked through during the session as an informal discussion, 
rather than set as exercises.

\subsection{Sessions}

List of the sessions in suggested reading order

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item \emph{Events}
\item \emph{Counting}
\item \emph{Conditioning}
\item \emph{Distributions}
\item \emph{Inequalities}
\item \emph{Simulation}
\item \emph{Intervals}
\end{enumerate}

\end{document}
